---
title: "CEE218X Assignment 1"
author: "Stephanie Gady"
date: "9/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
{r}
install.packages("tigris")
install.packages("sf")
install.packages("mapview")
install.packages("leaflet")
install.packages("censusapi")

```{r}
knitr::opts_chunk$set(warning = F, message = F)
library(tigris)
library(tidyverse)
library(sf)
library(mapview)
library(leaflet)
library(censusapi)
Sys.setenv(CENSUS_KEY="c8aa67e4086b4b5ce3a8717f59faa9a28f611dab")
```
```{r}
years <- 2017:2021
quarters <- 1:4
types <- c("Electric","Gas")
pge_data_raw <- NULL
for(year in years) {
  for(quarter in quarters) {
    for(type in types) {
      
      filename <- 
        paste0(
          "pge/PGE_",
          year,
          "_Q",
          quarter,
          "_",
          type,
          "UsageByZip.csv"
        )
  
      if (year == 2021&quarter %in% 3:4){
        next
        }
      
      temp <- read_csv(filename)
      
      if(type == "Electric") {
        temp <-
          temp %>%
          mutate(TOTALKBTU = TOTALKWH * 3.41214) %>%
          select(-TOTALKWH, -AVERAGEKWH)
      }
      if(type == "Gas") {
        temp <-
          temp %>%
          mutate(TOTALKBTU = TOTALTHM * 100) %>%
          select(-TOTALTHM, -AVERAGETHM)
      }
      
      pge_data_raw <- 
        rbind(pge_data_raw,temp)
      
    }
  }
}
```

```{r}
pge_data <-
  pge_data_raw %>% 
  filter(
    CUSTOMERCLASS %in% c("Elec- Commercial","Elec- Residential", "Gas- Commercial", "Gas- Residential")
  ) %>% 
  group_by(
    MONTH, 
    YEAR, 
    CUSTOMERCLASS
  ) %>% 
  summarize(
    TOTALKBTU = sum(TOTALKBTU, na.rm = T)
  ) %>% 
  mutate(
    DATE = 
      paste(
        YEAR,
        MONTH, 
        "01",
        sep="-"
      ) %>% as.Date()
  )
```

```{r}
pge_chart_res <-
  pge_data %>% 
  filter(CUSTOMERCLASS %in% c("Elec- Residential", "Gas- Residential")) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = MONTH %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTUs",
    title = "Residential Usage",
    fill = "Electricity Type"
  )

pge_chart_res
```
The above chart shows the average residential usage of electricity and gas by month. As you can see, the total energy used generally decreases in the spring and summer and begins to increase again in the fall and winter when heating is required. This difference may have been exhasperated by the pandemic as well, with more people spending the majority of the day at home. 
```{r}
pge_chart_commercial <-
  pge_data %>% 
  filter(CUSTOMERCLASS %in% c("Elec- Commercial", "Gas- Commercial")) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = MONTH %>% factor(),
      y = TOTALKBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTUs",
    title = "Commercial Usage",
    fill = "Electricity Type"
  )
pge_chart_commercial
```
The above chart shows the average commercial usage of electricity and gas by month. As you can see, the total energy used generally decreases in the spring and begins to increase again in the fall, similar to the residential plot. However, this change is less drastic, perhaps because commercial buildings rely less on natural ventilation. Overall commercial usage has likely decreased due to the pandemic, but is probably on the rise again as people return to offices in most of the country. 